Industrial, large-scale model predictive control with structured neural networks

### 2020-11-22 paper.tex
incorporate SJW's changes and add abstract/coverletter.

\begin{abstract}
The design of neural networks (NNs) is presented for treating large, linear
model predictive control (MPC) applications that are out of reach with available
quadratic programming (QP) solvers. First, we introduce a new feedforward
network architecture that enables practitioners to obtain offset-free
closed-loop performance with NNs. Second, we discuss the data generation
procedure to sample the relevant state space for the network training based on
anticipated online setpoint changes and plant disturbances. Third, we use the
input-to-state stability results available in the MPC literature and establish
robustness properties of NN controllers. Finally, we present illustrative
simulation studies on process control examples. We apply the NN design approach
and compare the performance with online QP based MPC on an industrial crude
distillation unit model with 252 states, 32 control inputs, and a control-sample
horizon length of 140. Parallel computing is used for data generation and
graphical processing units are used for network training. All the anticipated
plant operational scenarios with setpoints and disturbances that may change
during operation must be sampled for NN training. After the offline design
phase, NNs execute MPC three to five orders of magnitude faster than an
available QP solver with less than $1\%$ loss in the closed-loop performance.
\end{abstract}

@@

Model predictive control (MPC) is an online optimization based feedback control
technology. A dynamic model of the plant is used to make forecasts of the plant
measurements in response to the actuator movements, and an optimization is
solved in real time to determine the optimal actuator move to be applied to the
plant. For linear plant models, the optimization to be solved online is a
quadratic program (QP). The development of efficient QP algorithms
\citep*{kouzoupis:frison:zanelli:diehl:2018, wright:2019} allowed practitioners
to apply MPC in the process industries \citep*{qin:badgwell:2003, lahiri:2017}.

@@

To make progress in the size of control applications achievable with MPC, large
problems in which QP solvers fail to deliver the control in the available sample
time should be addressed. The scalability of the NN approach to such problems
with large model dimensions and control horizon length in the QP has not been
demonstrated. Two issues arise in these problems: i) In a large dimensional
state space, the entire set of feasible states of the QP cannot be sampled
densely for the NN training. This issue is a form of the well-known curse of
dimensionality. ii) One may sample the state space partially to avoid this
problem, but the time required for the data generation can still be impractical
due to the long time required to solve a QP for each sampled state. We
demonstrate in this article by case studies that this latter issue can be
resolved for a large size of problems with parallel computing, and the former
issue by sampling the state space for typical plant operational scenarios. The
following feature must be present in the large-scale application of interest for
the partial sampling scheme proposed in this article to be applicable: for a
particular operating mode of the plant, the number of frequently changing
setpoints and large magnitude disturbances must be reasonably small such that
the entire large dimensional state space is not visited during that operation
mode. The MPC feedback law can then be reliably approximated using NNs for the
different modes of plant operations driven by their respective set of setpoint
changes and disturbances.

From the above viewpoint, this article discusses the design of NNs as an
alternative to real time optimization in MPC, with an emphasis on large
applications which are challenging with QP solvers. The focus of this article is
on setpoint tracking MPC problems relevant for deployment in the process
industries. We present a new feedforward network architecture to obtain
offset-free closed-loop performance with the NNs. Additionally, we use the
input-to-state stability results available in the MPC literature, and establish
conditions under which the NN controller is robust to state estimation errors
and process disturbances. We present case studies on process control examples
that demonstrate the scalability of the proposed NN design approach.

@@

\subsection{Target selector}
Given the disturbance estimate, the input and controlled measurement setpoints,
we consider the following target selector QP
\begin{align}
    \min_{x_s, \ u_s}  \dfrac{1}{2} |u_{sp} - u_s|^2_{R_s}
\end{align}
subject to
\begin{align}  
    \begin{bmatrix}
        I -A & -B \\
        HC & 0
    \end{bmatrix} \begin{bmatrix}
        x_s \\
        u_s
    \end{bmatrix} &= \begin{bmatrix}
        B_d\dhat \\
        r_{sp} - HC_d\dhat
    \end{bmatrix} \\
    \underline{u} \leq u_s \leq \overline{u}
\end{align}
in which ($x_s, u_s$) is the target steady-state pair, $r=Hy$ are the controlled
measurements chosen as some subset or a linear combination of all the
measurements, ($\underline{u}, \overline{u}$) are the actuator bounds, and
($u_{sp}, r_{sp}$) are the input and controlled measurement setpoints. The
equality constraint $r_{sp} = H(Cx_s + C_d\dhat)$ can be difficult to satisfy
exactly in real time depending on the controlled measurement setpoint and
disturbance estimate. Then, the hard setpoint equality constraint can be relaxed
and moved to the stage cost such that the target selector computes a steady
state to minimize the offset in the controlled measurements. We assume for the
simulation studies in this article that the input setpoint ($u_{sp}$) is fixed
at some chosen steady state and only the controlled measurement setpoint
($r_{sp}$) changes in real time.

@@

\subsection{Structured neural network}
An intuitive approach is to build a feedforward NN that takes the triple ($x,
x_s, u_s$) as its input and outputs a control that is close to the optimal
control. This strategy has been proposed in \cite*{karg:lucia:2020} and 
\cite*{chen:saulnier:atanasov:lee:kumar:pappas:morari:2018}
for a single fixed steady state. The approach has not been demonstrated to scale
to the size of model dimensions and the setpoint tracking offset-free MPC
problem considered in this article. An issue with this strategy is that the NN
has no knowledge about the MPC feedback law, and can require large amounts of
training data in large state dimensions to obtain a reasonable controller. At a
minimum, the MPC feedback law is such that at the origin, the control action
taken by the MPC controller is zero. For the setpoint tracking MPC problem
considered in this article, the NN control law must be such that at steady-state
operations, the NN maintains the plant at the desired steady state such that the
controlled measurements are at their setpoints. This structure implies $u_s =
\kappa_{NN}(x = x_s, x_s, u_s)$, in which we use $\kappa_{NN}(\cdot)$ to denote
the NN control law. To incorporate this information in a NN, we introduce the
following architecture
\begin{align}
z_0 &= \begin{bmatrix}
    x', & x_s', & u_s', & x_s', & x_s', & u_s' \end{bmatrix}'\notag \\
    f_i(z_{i-1}) &= \begin{bmatrix}
        W_i & 0 \notag \\
        0 & W_i
          \end{bmatrix}z_{i-1}  + \begin{bmatrix}
            b_i \notag 
            b_i
            \end{bmatrix} \notag\\
    z_i &= g(f_i(z_{i-1})), \textnormal{for} \ i \in \bbI_{1:h} \notag \\
    u &= u_s + \begin{bmatrix}
  W_{h + 1}, & -W_{h + 1} \\
    \end{bmatrix} z_h
\label{eq:NNstruct}
\end{align}
in which $g(a) = \textnormal{max} (0, a)$, is the ReLU operation applied
element-wise on its vector input, $z_i$ is the output of each hidden layer of
the network, $z_0$ is the input to the network, $h$ is the number of hidden
layers, and $i \in \bbI_{1:h}$. The parameters to be optimized using training
data are ($W_i, b_i$) and $W_{h+1}$, which are matrices of appropriate
dimensions. At steady state, when $x=x_s$ in $z_0$, \eqref{eq:NNstruct} outputs
$u_s$ regardless of the choice of weights. The software tensorflow
\citep{abadi:agarwal:barham:brevdo:et-al:2015} is used for NN training purposes
in this article, and the structured architecture can be built with symbolic
differentiation tools available in tensorflow. When a rate-of-change penalty on
the control is included in the regulator QP, the input to the structured NN is
modified to 
\begin{align*}
    z_0 &= 
    \begin{bmatrix}
    x', & u_{-1}', & x_s', & u_s', & x_s', & u_s', & x_s', & u_s' \\
    \end{bmatrix}'
\end{align*}
in which $u_{-1}$ is the previous control applied to the plant and the
dimensions of the matrices ($W_i, b_i$) and $W_{h+1}$ are appropriately
adjusted. Even after training, the control produced by the NN may not satisfy
the hard actuator constraints, and the saturation function
($\textnormal{sat}(\cdot, \underline{u}, \overline{u})$) is applied to the
output of the NN.

@@

For one fixed steady state, the set of feasible states for the MPC QP defined by
\eqref{eqn:regulator}-\eqref{eqn:regulator_constraints} is the entire
$n$-dimensional state space $\bbR^n$. Additionally, the target steady state may
change during the plant operation depending on the setpoint and disturbance
estimate. The domain of the MPC control law ($\kappa_N(x, x_s, u_s)$) is
therefore the state space $\bbR^{n}$ and the possible set of steady-state target
pairs. Sampling this entire domain of the MPC control law for NN training is
difficult in large dimension, and we show in Section
\ref{sec:simulation_studies} that it is also not required for practical
applications.

From a plant operations perspective, a majority of the large-scale petrochemical
plants operate in a relatively small number of operating regimes or scenarios.
Each operational scenario is driven by a selected few controlled measurement
setpoints that depend on product demands and some large magnitude disturbances
may change frequently, while the setpoints for the other measurements remain
constant for long periods of time. The states and steady-state targets spanned
in a closed-loop operation per scenario is thereby considerably less than the
entire domain of the MPC control law. This typical operation of large chemical
plants can be exploited, and we propose the following data generation procedure
to perform offline simulations using the model to collect the operationally
relevant set of ($x, x_s, u_s, \kappa_N(\cdot)$) for the NN training:

@@

\begin{figure*}[!h]
    \centering
	\includegraphics[page=1, width=0.7\textwidth,
		height=0.5\textheight]{cstrs_comparision_plots.pdf}    
        \caption{Transient closed-loop actuator trajectories of the CSTRs with
        flash separator example with the NN and optimal MPC controllers.}   
    \label{fig:cl_cstrs_inputs}
\end{figure*}

\begin{figure*}[!h]
    \centering
	\includegraphics[page=2, width=0.7\textwidth,
		height=0.5\textheight]{cstrs_comparision_plots.pdf} \caption{Transient
		closed-loop controlled measurement trajectories of the CSTRs with flash
		separator example with the NN and optimal MPC controllers.}
		\label{fig:cl_cstrs_outputs}
\end{figure*}

The robustness property of NN controllers is now presented. The optimal MPC
controller designed for a nominal linear system ignoring disturbances is known
to be inherently robust to bounded state estimation errors and disturbances
\citep*{heath:wills:2005, pannocchia:rawlings:wright:2011}. The approximation
error of the MPC feedback law by a NN can be viewed as an additional
disturbance, and the existing input-to-state stability results
\citep*{sontag:wang:1995} are used to establish the allowable approximation
error in the NN for the NN controller to be robust to disturbances. Theorem
\ref{thm:nnrobustness} states this allowable approximation error and the size of
disturbances for which the NN controller is robust.

The problem of steering the state of the linear system (\ref{eq:ltimodel}) to
one fixed steady state ($x_s, u_s$) is considered. For the analysis in this
section, the state and control are transformed in deviation variables as $x := x
- x_s$ and $u := u -u_s$, and define the control problem as regulation to the
origin. The optimal MPC control law is denoted as $\kappa_N(x)$, and the NN
control law as $\kappa_{NN}(x) = \kappa_N(x) + e_{NN}(x)$, in which $e_{NN}(x)$
is the approximation error of the optimal control law by the network. Assume in
the analysis that the target steady state is such that the unconstrained linear
quadratic regulator (LQR) satisfies the actuator constraints in a small
neighborhood near the origin.

As reviewed in \cite*{mayne:rawlings:rao:scokaert:2000}, the nominal stability
of MPC can be established using the optimal cost function ($V_N^0(x)$) of the QP
as a Lyapunov function. The region of attraction of the closed-loop system $x^+
= Ax + B\kappa_N(x)$ when no hard terminal region constraint is included in the
QP can be characterized as $\mathcal{X}_N = \textnormal{lev}_{Nd + \tau} \
V_N^0(x)$ \citep*{limon:alamo:salas:camacho:2006}. The parameter $\tau$ is the
level set parameter of the terminal region chosen as $\bbX_f =
\textnormal{lev}_{\tau} \ x'Px$, and $d > 0$ is a constant such that $\ell(x, u)
\geq d$ for all $x \in \bbR^n \setminus \bbX_f$ and $\underline{u} \leq u+u_s
\leq \overline{u}$. The parameter $\tau$ is chosen small enough such that the
unconstrained LQR satisfies the actuator constraints for all $x \in \bbX_f$. The
optimal cost function is continuous and satisfies $c_1\norm{x}^2 \leq V_N^0(x)
\leq c_2\norm{x}^2$ and $V_N^0(x^+) - V_N^0(x) \leq -c_1\norm{x}^2$ for some
$c_2 \geq c_1 > 0$.

@@

The proof of this theorem along with the required input-to-state stability
definitions are provided in Appendix \ref{app:theorem1}. In the nominal case,
the disturbance sequences $\mathbf{e}$ and $\mathbf{w}$ are zero. The
closed-loop states in this case have the bound $\norm{\phi(k; x)} \leq
\beta(\norm{x}, k) + \alpha_n(\bar{e}_{NN})$. Any additional value of the NN
training is to reduce the worst approximation error ($\bar{e}_{NN}$) in the
state space of interest and reduce the bound on the closed-loop states.

@@

In this section, two case studies are presented that demonstrate the scalability
of the proposed NN design approach. The closed-loop performance of NNs and
computational benefits over online QP based MPC are analyzed in these case
studies. After the offline design of NNs, two types of validation simulations
are performed: (i) with setpoints and disturbances not present in the training
data but within the same range of values used to generate the training data,
(ii) with extra setpoint signals assumed to be unknown during the offline
network design phase. Both these types of simulations are conducted to analyze
the interpolation and extrapolation abilities of the NNs, and are referred
subsequently as simulations with \textit{expected} and \textit{unexpected} plant
behaviors, respectively.

@@

\begin{figure}[h]
    \centering
    \includegraphics[width=0.5\textwidth,
                     height=0.22\textheight]{cstrs_with_flash.pdf}
    \caption{Schematic of the CSTRs with a flash separator model}
    \label{fig:schematic_cstrs}
\end{figure}

@@

\begin{table*}[t]
    \caption{Summary of the simulation study for the CSTRs with a flash
    separator example.}   
    \begin{tabular}{ |c|c|c|c|c|c|c|c| }
        \hline
        Controller & \thead{Structured network \\ architecture} & \thead{Number
        of \\ Parameters} & \thead{Memory \\ footprint \\ (mB)} &
        \thead{Training \\ time \\ (min)} & \thead{\% Performance\\ losses \\
        Expected, Unexpected \\ plant behaviors} & \thead{Average \\ speedup} &
        \thead{Worst case \\ speedup} \\
        \hline
    SS &  &  & &  & 85.18\%, 106.39 \% &  & \\ 
    satK &  &  &  &  & 41.03\%, 27.61 \% & $2.47 \times 10^5$ & $3.07 \times
    10^5$ \\ 
    SH &  &  &  &  & 1.61\%, 2.46 \% & $3.75 \times 10^3$ & $8.4 \times 10^1$ \\

    NN-UNS &  &  &  &  & 80.49\%, 73.29 \% & $9.31 \times 10^4$ & $4.99 \times
    10^3$ \\ 
NN-3-448 & [72, 448, 448, 448, 6] & $1.10 \times 10^5$ & 0.84 & 13.02 & 0.28 \%,
5.57 \% & $5.26 \times 10^4$ & $1.45 \times 10^3$ \\ 
NN-3-480 & [72, 480, 480, 480, 6]  & $1.26 \times 10^5$ & 0.96  & 12.51 & 0.34
\%, 7.58\% & $5.09 \times 10^4$ & $1.53 \times 10^4$ \\ 
NN-3-512 & [72, 512, 512, 512, 6]  & $1.42 \times 10^5$ & 1.08  & 12.77 & 0.16
\%, 5.90 \% & $4.62 \times 10^4$ & $1.58 \times 10^4$ \\ 
        \hline
       \end{tabular}
       \label{table:cstrs}      
\end{table*}

\begin{figure*}[!h]
    \includegraphics[page=31, width=0.5\textwidth,
        height=0.28\textheight]{cstrs_comparision_plots.pdf} \hfill
        \includegraphics[page=34, width=0.5\textwidth,
        height=0.28\textheight]{cstrs_comparision_plots.pdf} \vfill
    \vspace{-0.1in}
    \begin{center}\includegraphics[page=33, width=0.5\textwidth,
        height=0.28\textheight]{cstrs_comparision_plots.pdf}
    \end{center}
    \vspace{-0.2in}
    \caption{Simulation statistics for the  
    CSTRs with a flash separator example. Data requirement to obtain accurate NN
    controllers (top left), online computation times of the NNs and MPC
    controller (top right), and the controller performance index obtained with
    the NNs, SS, satK and MPC controllers (center bottom).}	
    \label{fig:cstrs_statistics}
\end{figure*}

@@

The $\%$ performance loss obtained using the four heuristic controllers, their
corresponding speedups, maximum time required to train each NN, number of
parameters in each NN, and the memory required to store  
the weights of the NNs is summarized in the Table \ref{table:cstrs}. The $85 \%$
performance loss reported for the unstructured NN corresponds to training with
$1.5 \times 10^5$ samples and the performance is unacceptable for an industrial
application. The noticeable performance improvement with the structured NNs is
because the control law is exactly the same as the optimal feedback law at
steady-states, and that simplifies the optimal feedback law approximation
problem since the training process only has to improve the NN control law for
the transient conditions.

@@

Four structured NN architectures are considered for this example. Each NN has
three hidden layers, with 1664 (NN-3-1664), 1792 (NN-3-1792), 1920 (NN-3-1920),
and 2048 (NN-3-2048) nodes in each hidden layer respectively. For the validation
simulations, one set of setpoint and disturbance signals for the expected plant
behavior case is generated with 24 setpoint changes and 48 disturbance changes
for 2880 timesteps (2 days). The short horizon MPC controller is constructed
with a horizon length of $N = 3$. The states and target steady-states ($x$,
$x_s$) are scaled for the NN training as in the previous example.

>>>>

\begin{abstract}
The design of neural networks (NNs) is presented for treating large, linear
model predictive control (MPC) applications that are out of reach with available
quadratic programming (QP) solvers. First, we introduce a new feedforward
network architecture that enables practitioners to obtain offset-free
closed-loop performance with NNs. Second, we discuss the data generation
procedure to sample the state space relevant to training the NNs based on
anticipated online setpoint changes and plant disturbances. Third, we use the
input-to-state stability results available in the MPC literature and establish
robustness properties of NN controllers. Finally, we present illustrative
simulation studies on process control examples. We apply the NN design approach
and compare the performance with online QP based MPC on an industrial crude
distillation unit model with 252 states, 32 control inputs, and a control-sample
horizon length of 140. Parallel computing is used for data generation and
graphical processing units are used for network training. All the anticipated
plant operational scenarios with setpoints and disturbances that may change
during operation must be sampled for NN training. After the offline design
phase, NNs execute MPC three to five orders of magnitude faster than an
available QP solver with less than $1\%$ loss in the closed-loop performance.
\end{abstract}

@@

Model predictive control (MPC) is an online optimization based feedback control
technology. A dynamic model of the plant is used to make forecasts of the plant
measurements in response to the actuator movements, and an optimization problem
is solved in real time to determine the optimal actuator move to be applied to
the plant. For linear plant models, the optimization problem is a quadratic
program (QP). The development of efficient QP algorithms
\citep*{kouzoupis:frison:zanelli:diehl:2018, wright:2019} allowed practitioners
to apply MPC in the process industries \citep*{qin:badgwell:2003, lahiri:2017}.

@@

To make progress in the size of control applications achievable with MPC, large
problems in which QP solvers fail to deliver the control in the available sample
time should be addressed. The scalability of the NN approach to such problems
with large model dimensions and control horizon length in the QP has not been
demonstrated to date. Two issues arise in these problems: i) In a large
dimensional state space, the entire set of feasible states of the QP cannot be
sampled densely for the NN training. This issue is a form of the well-known
curse of dimensionality. ii) One may sample the state space partially to avoid
this problem, but the time required for the data generation can still be
impractical due to the long time required to solve a QP for each sampled state.
We demonstrate in this article by case studies that this latter issue can be
resolved for a large size of problems with parallel computing, and the former
issue by sampling the state space for typical plant operational scenarios. The
following feature must be present in the large-scale application of interest for
the partial sampling scheme proposed in this article to be applicable: for a
particular operating mode of the plant, the number of frequently changing
setpoints and large magnitude disturbances must be reasonably small such that
the entire large dimensional state space is not visited during that operation
mode. The MPC feedback law can then be reliably approximated using NNs for the
different modes of plant operations driven by their respective set of setpoint
changes and disturbances.

From the above viewpoint, this article discusses the design of NNs as an
alternative to real time optimization in MPC, with an emphasis on large
applications that are challenging for QP solvers. The focus of this article is
on setpoint tracking MPC problems relevant for deployment in the process
industries. We present a new feedforward network architecture to obtain
offset-free closed-loop performance with the NNs. Additionally, we use the
input-to-state stability results available in the MPC literature, and establish
conditions under which NN controllers are robust to state estimation errors
and process disturbances. We present case studies on process control examples
that demonstrate the scalability of the proposed NN design approach.

@@

\subsection{Target selector}
Given the disturbance estimate, the input and controlled measurement setpoints,
we consider the following target selector QP
\begin{align}
    \min_{x_s, \ u_s}  \dfrac{1}{2} |u_{sp} - u_s|^2_{R_s}
\end{align}
subject to
\begin{align}  
    \begin{bmatrix}
        I -A & -B \\
        HC & 0
    \end{bmatrix} \begin{bmatrix}
        x_s \\
        u_s
    \end{bmatrix} &= \begin{bmatrix}
        B_d\dhat \\
        r_{sp} - HC_d\dhat
    \end{bmatrix} \\
    \underline{u} \leq u_s \leq \overline{u}
\end{align}
in which ($x_s, u_s$) is the target steady-state pair, $r=Hy$ are the controlled
measurements chosen as some subset or a linear combination of all the
measurements, ($\underline{u}, \overline{u}$) are the actuator bounds, and
($u_{sp}, r_{sp}$) are the input and controlled measurement setpoints. The
equality constraint $r_{sp} = H(Cx_s + C_d\dhat)$ can be difficult to satisfy
exactly in real time depending on the controlled measurement setpoint and
disturbance estimate. In this case, the hard setpoint equality constraint can be
relaxed and moved to the stage cost such that the target selector computes a
steady state to minimize the offset in the controlled measurements. We assume
for the simulation studies in this article that the input setpoint ($u_{sp}$) is
fixed at some chosen steady state and only the controlled measurement setpoint
($r_{sp}$) changes in real time.

@@

\subsection{Structured neural network}
An intuitive approach is to build a feedforward NN that takes the triple ($x,
x_s, u_s$) as its input and outputs a control that is close to the optimal
control. This strategy has been proposed in \cite*{karg:lucia:2020} and 
\cite*{chen:saulnier:atanasov:lee:kumar:pappas:morari:2018}
for a single fixed steady state. The approach has not been demonstrated to scale
to the size of model dimensions and the setpoint tracking offset-free MPC
problem considered in this article. An issue with this strategy is that the NN
has no knowledge about the MPC feedback law, and can require large amounts of
training data in large state dimensions to obtain a reasonable controller. At a
minimum, the MPC feedback law is such that at the origin, the control action
taken by the MPC controller is zero. For the setpoint tracking MPC problem
considered in this article, the NN control law must be such that at steady-state
operations, the NN maintains the plant at the desired steady state such that the
controlled measurements are at their setpoints. This structure implies $u_s =
\kappa_{NN}(x = x_s, x_s, u_s)$, in which we use $\kappa_{NN}(\cdot)$ to denote
the NN control law. To incorporate this information in a NN, we introduce the
following architecture
\begin{align}
z_0 &= \begin{bmatrix}
    x', & x_s', & u_s', & x_s', & x_s', & u_s' \end{bmatrix}'\notag \\
    f_i(z_{i-1}) &= \begin{bmatrix}
        W_i & 0 \notag \\
        0 & W_i
          \end{bmatrix}z_{i-1}  + \begin{bmatrix}
            b_i \\
            b_i
            \end{bmatrix} \notag\\
    z_i &= g(f_i(z_{i-1})), \textnormal{for} \ i \in \bbI_{1:h} \notag \\
    u &= u_s + \begin{bmatrix}
  W_{h + 1}, & -W_{h + 1} \\
    \end{bmatrix} z_h
\label{eq:NNstruct}
\end{align}
in which $g(a) = \textnormal{max} (0, a)$, is the ReLU operation applied
element-wise on its vector input, $z_i$ is the output of each hidden layer of
the network, $z_0$ is the input to the network, $h$ is the number of hidden
layers, and $i \in \bbI_{1:h}$. The parameters to be optimized using training
data are ($W_i, b_i$) and $W_{h+1}$, which are matrices of appropriate
dimensions. At steady state, when $x=x_s$ in $z_0$, \eqref{eq:NNstruct} outputs
$u_s$ regardless of the choice of weights. The software tensorflow
\citep{abadi:agarwal:barham:brevdo:et-al:2015} is used for NN training purposes
in this article, and the structured architecture can be built with symbolic
differentiation tools available in tensorflow. When a rate-of-change penalty on
the control is included in the regulator QP, the input to the structured NN is
modified to 
\begin{align*}
    z_0 &= 
    \begin{bmatrix}
    x', & u_{-1}', & x_s', & u_s', & x_s', & u_s', & x_s', & u_s' \\
    \end{bmatrix}'
\end{align*}
in which $u_{-1}$ is the previous control applied to the plant and the
dimensions of the matrices ($W_i, b_i$) and $W_{h+1}$ are appropriately
adjusted. Even after training, the control produced by the NN may not satisfy
the hard actuator constraints, and the saturation function
($\textnormal{sat}(\cdot, \underline{u}, \overline{u})$) is applied to the
output of the NN.

@@

For one fixed steady state, the set of feasible states for the MPC QP defined by
\eqref{eqn:regulator}-\eqref{eqn:regulator_constraints} is the entire
$n$-dimensional state space $\bbR^n$. Additionally, the target steady state may
change during the plant operation depending on the setpoint and disturbance
estimate. The domain of the MPC control law ($\kappa_N(x, x_s, u_s)$) is
therefore the state space $\bbR^{n}$ and the possible set of steady-state target
pairs. Sampling this entire domain of the MPC control law for NN training is
difficult in large dimensions, and we show in Section
\ref{sec:simulation_studies} that it is also not required for practical
applications.

From a plant operations perspective, a majority of the large-scale petrochemical
plants operate in a relatively small number of operating regimes or scenarios.
Each operational scenario is driven by a selected few controlled measurement
setpoints that depend on product demands and some large magnitude disturbances
may change frequently, while the setpoints for the other measurements remain
constant for long periods of time. The states and steady-state targets spanned
in a closed-loop operation per scenario is thereby considerably less than the
entire domain of the MPC control law. This typical operation of large chemical
plants can be exploited, and we propose the following data generation procedure
to perform offline simulations using the model to collect the operationally
relevant set of ($x, x_s, u_s, \kappa_N(\cdot)$) for NN training.

@@

\begin{figure*}[!h]
    \centering
	\includegraphics[page=1, width=0.7\textwidth,
		height=0.5\textheight]{cstrs_comparision_plots.pdf}    
        \caption{Transient closed-loop actuator trajectories of the CSTRs with
        flash separator example with the NN and optimal MPC controllers. The
        performance of NN and optimal MPC controllers are almost
        indistinguishable.}   
    \label{fig:cl_cstrs_inputs}
\end{figure*}

\begin{figure*}[!h]
    \centering
	\includegraphics[page=2, width=0.7\textwidth,
		height=0.5\textheight]{cstrs_comparision_plots.pdf} \caption{Transient
		closed-loop controlled measurement trajectories of the CSTRs with flash
		separator example with the NN and optimal MPC controllers. The
		performance of NN and optimal MPC controllers are almost
		indistinguishable.}
		\label{fig:cl_cstrs_outputs}
\end{figure*}

We now present the robustness properties of NN controllers. The optimal MPC
controller designed for a nominal linear system ignoring disturbances is known
to be inherently robust to bounded state estimation errors and disturbances
\citep*{heath:wills:2005, pannocchia:rawlings:wright:2011}. The approximation
error of the MPC feedback law by a NN can be viewed as an additional
disturbance, and the existing input-to-state stability results
\citep*{sontag:wang:1995} are used to establish the allowable approximation
error in the NN for the NN controller to be robust to disturbances. Theorem
\ref{thm:nnrobustness} states this allowable approximation error and specifies
the size of disturbances for which the NN controller is robust.

The problem of steering the state of the linear system (\ref{eq:ltimodel}) to
one fixed steady state ($x_s, u_s$) is considered. For the analysis in this
section, the state and control are transformed in deviation variables as $x := x
- x_s$ and $u := u -u_s$, and we define the control problem as regulation to the
origin. The optimal MPC control law is denoted as $\kappa_N(x)$, and the NN
control law as $\kappa_{NN}(x) = \kappa_N(x) + e_{NN}(x)$, in which $e_{NN}(x)$
is the approximation error of the optimal control law by the network. Assume in
the analysis that the target steady state is such that the unconstrained linear
quadratic regulator (LQR) satisfies the actuator constraints in a small
neighborhood near the origin.

As reviewed in \cite*{mayne:rawlings:rao:scokaert:2000}, the nominal stability
of MPC can be established using the optimal cost function ($V_N^0(x)$) of the QP
as a Lyapunov function. The region of attraction of the closed-loop system $x^+
= Ax + B\kappa_N(x)$ when no hard terminal region constraint is included in the
QP can be characterized as $\mathcal{X}_N = \textnormal{lev}_{Nd + \tau} \
V_N^0(x)$ \citep*{limon:alamo:salas:camacho:2006}. The parameter $\tau$ is the
level set parameter of the terminal region chosen as $\bbX_f =
\textnormal{lev}_{\tau} \ x'Px$, $d > 0$ is a constant such that $\ell(x, u)
\geq d$ for all $x \in \bbR^n \setminus \bbX_f$, and $\underline{u} \leq u+u_s
\leq \overline{u}$. The parameter $\tau$ is chosen small enough such that the
unconstrained LQR satisfies the actuator constraints for all $x \in \bbX_f$. The
optimal cost function is continuous and satisfies $c_1\norm{x}^2 \leq V_N^0(x)
\leq c_2\norm{x}^2$ and $V_N^0(x^+) - V_N^0(x) \leq -c_1\norm{x}^2$ for some
$c_2 \geq c_1 > 0$.

@@

The proof of this theorem along with the required input-to-state stability
definitions are provided in Appendix \ref{app:theorem1}. In the nominal case,
the disturbance sequences $\mathbf{e}$ and $\mathbf{w}$ are zero. The
closed-loop states in this case have the bound $\norm{\phi(k; x)} \leq
\beta(\norm{x}, k) + \alpha_n(\bar{e}_{NN})$. Any additional value of the NN
training is to reduce the worst approximation error ($\bar{e}_{NN}$) in the
state space of interest and thus reduce the bound on the closed-loop states.

@@

In this section, two case studies are presented that demonstrate the scalability
of the proposed NN design approach. The closed-loop performance of NNs and
computational benefits over online QP based MPC are analyzed in these case
studies. After the offline design of NNs, two types of validation simulations
are performed: (i) with setpoints and disturbances not present in the training
data but within the same range of values used to generate the training data,
(ii) with extra setpoint signals assumed to be unknown during the offline
network design phase. Both these types of simulations shed light on 
the interpolation and extrapolation abilities of the NNs, and are referred
subsequently as simulations with \textit{expected} and \textit{unexpected} plant
behaviors, respectively.

@@

\begin{figure}[h]
    \centering
    \includegraphics[width=0.5\textwidth,
                     height=0.22\textheight]{cstrs_with_flash.pdf}
    \caption{Schematic of the CSTRs with a flash separator model.}
    \label{fig:schematic_cstrs}
\end{figure}

@@

\begin{table*}[t]
    \caption{Summary of the simulation study for the CSTRs with a flash
    separator example.}   
    \begin{tabular}{ |c|c|c|c|c|c|c|c| }
        \hline
        Controller & \thead{Neural network \\ architecture} & \thead{Number of
        \\ Parameters} & \thead{Memory \\ footprint \\ (mB)} & \thead{Training
        \\ time \\ (min)} & \thead{\% Performance\\ losses \\
        Expected, Unexpected \\ plant behaviors} & \thead{Average \\ speedup} &
        \thead{Worst case \\ speedup} \\
        \hline
    SS &  &  & &  & 85.18\%, 106.39 \% &  & \\ 
    satK &  &  &  &  & 41.03\%, 27.61 \% & $2.47 \times 10^5$ & $3.07 \times
    10^5$ \\ 
    SH &  &  &  &  & 1.61\%, 2.46 \% & $3.75 \times 10^3$ & $8.4 \times 10^1$ \\

    NN-UNS & [36, 224, 224, 224, 6] & $1.10 \times 10^5$ &  &  & 80.49\%, 73.29 \% & $9.31 \times
    10^4$ & $4.99 \times 10^3$ \\ 
NN-3-448 & [72, 448, 448, 448, 6] & $1.10 \times 10^5$ & 0.84 & 13.02 & 0.28 \%,
5.57 \% & $5.26 \times 10^4$ & $1.45 \times 10^3$ \\ 
NN-3-480 & [72, 480, 480, 480, 6]  & $1.26 \times 10^5$ & 0.96  & 12.51 & 0.34
\%, 7.58\% & $5.09 \times 10^4$ & $1.53 \times 10^4$ \\ 
NN-3-512 & [72, 512, 512, 512, 6]  & $1.42 \times 10^5$ & 1.08  & 12.77 & 0.16
\%, 5.90 \% & $4.62 \times 10^4$ & $1.58 \times 10^4$ \\ 
        \hline
       \end{tabular}
       \label{table:cstrs}      
\end{table*}

\begin{figure*}[!h]
    \includegraphics[page=31, width=0.5\textwidth,
        height=0.28\textheight]{cstrs_comparision_plots.pdf} \hfill
        \includegraphics[page=34, width=0.5\textwidth,
        height=0.28\textheight]{cstrs_comparision_plots.pdf} \vfill
    \vspace{-0.1in}
    \begin{center}\includegraphics[page=33, width=0.5\textwidth,
        height=0.28\textheight]{cstrs_comparision_plots.pdf}
    \end{center}
    \vspace{-0.2in}
    \caption{Simulation statistics for the  
    CSTRs with a flash separator example. Data requirement to obtain accurate NN
    controllers (top left), online computation times of the NNs and MPC
    controller (top right), and the controller performance index obtained with
    the NNs, SS, satK and MPC controllers (center bottom).The performance of the
    three NN and optimal MPC controllers are indistinguishable.}	
    \label{fig:cstrs_statistics}
\end{figure*}

@@

The $\%$ performance loss obtained using the four heuristic controllers, their
corresponding speedups, maximum time required to train each NN, number of
parameters in each NN, and the memory required to store  
the weights of the NNs is summarized in the Table \ref{table:cstrs}. The $85 \%$
performance loss reported for the unstructured NN corresponds to training with
$1.5 \times 10^5$ samples and the performance is unacceptable for an industrial
application. The noticeable performance improvement with the structured NNs is
because the control law is exactly the same as the optimal feedback law at
steady-states, which simplifies the optimal feedback law approximation
problem since the training process only has to improve the NN control law for
the transient conditions.

@@

Three structured NN architectures are considered for this example. Each NN has
three hidden layers, with 1664 (NN-3-1664), 1792 (NN-3-1792), and 1920
(NN-3-1920) nodes in each hidden layer respectively. For the validation
simulations, one set of setpoint and disturbance signals for the expected plant
behavior case is generated with 24 setpoint changes and 48 disturbance changes
for 2880 timesteps (2 days). The short horizon MPC controller is constructed
with a horizon length of $N = 3$. The states and target steady-states ($x$,
$x_s$) are scaled for the NN training as in the previous example.
